% ===========================================================================
% III. STRUCTURAL VERIFIABILITY OF NEURAL NETWORKS
% ===========================================================================
\section{Structural Verifiability of Neural Networks}
\label{sec:svnn}

We now formalize a central observation that emerged from the compiler design
process: whether a neural network compiler can provide design-time correctness
guarantees is not solely a property of the compiler---it is fundamentally
determined by the \textit{architecture} being compiled. This section defines
the class of architectures for which correctness-preserving compilation is
achievable, proves that KAN belongs to this class, and demonstrates that
standard MLPs do not.

\vspace{4pt}
\begin{remark}[The Compilable Frontier]
\label{rem:compilable_frontier}
The SVNN conditions define a \textbf{compilable frontier}---a principled
boundary separating neural architectures whose deployed behavior can be
certified at design time from those that require runtime testing. Architectures
satisfying Conditions~1--2 lie on the \textit{interior} (certifiable side):
their decomposition into pure linear and element-wise operations enables the
structural induction that Theorem~2's proof relies on. Architectures violating
any condition lie on the \textit{exterior} (empirical-only side): they can
still be compiled and deployed, but their correctness guarantees are limited
to test-set validation. Proposition~1 (below) establishes that standard MLPs
with SiLU, Sigmoid, or Tanh activations---the dominant architectures in
industrial ML---lie on the exterior. This frontier is not an implementation
artifact; it reflects a fundamental property of which function classes admit
closed-form, compiler-computable error propagation.
\end{remark}


\subsection{The Verification Challenge for PLC-Deployed Networks}
\label{sec:svnn-challenge}

Industrial PLCs govern safety-critical machinery: conveyor sorting systems,
motor drives, and process control loops where incorrect outputs can cause
equipment damage or production line stoppage. Deploying neural networks on
such controllers therefore demands correctness guarantees that go beyond
empirical test-set accuracy.

Existing ML$\to$PLC tools (RTNNIgen~\cite{rtnnigen2024}, MLconverter%
~\cite{mlconverter2025}, ICSML~\cite{icsml2023}) provide \textit{no}
mathematical guarantee that the generated PLC code preserves the trained
model's behavior. Their correctness model is purely empirical: train, convert,
and test on a held-out set. For industrial deployment---where inputs may
drift, sensors may degrade, and the PLC's REAL arithmetic differs from
PyTorch's float32---this is insufficient.

We identify three levels of deployment verification, in increasing order
of strength:

\begin{enumerate}
    \item \textbf{Level 1---Runtime Testing (Weakest):} Evaluate the deployed
          model on a test set and report accuracy. This is the status quo
          (RTNNIgen, MLconverter). It provides no guarantee on unseen inputs
          and cannot detect systematic floating-point discrepancies.
    \item \textbf{Level 2---Design-Time Error Bounds (SVNN):} Compute, at
          \textit{compile time} and without executing on hardware, a provable
          per-output error bound $\varepsilon_i$ such that for all valid
          inputs $x$, the deployed model's output differs from the trained
          model's output by at most $\varepsilon_i$. This is the guarantee
          that \neuroplc provides (Theorem~1).
    \item \textbf{Level 3---Mechanized Formal Verification (Strongest):}
          A machine-checked proof (e.g., in Coq, Isabelle, or Lean) that the
          compiler preserves semantics down to the IEEE~754 bit-level. This
          has been achieved for select neural network kernels in TorchLean%
          \cite{torchlean2025} but is currently intractable for full
          compiler pipelines.
\end{enumerate}

Level~2---design-time error bounds---is the appropriate level for
industrial deployment. The question is: \textit{which architectures admit
such bounds, and why?}

\subsection{Definition: Structurally Verifiable Neural Network (SVNN)}
\label{sec:svnn-definition}

\begin{definition}[Structurally Verifiable Neural Network]
\label{def:svnn}
A neural network architecture $\mathcal{A}$ is \textbf{structurally
verifiable} (SVNN) with respect to a compiler $C$ if for any instance
$M \in \mathcal{A}$, the compiler can compute, at design time and without
executing $M$ on target hardware, a per-output error bound
$\varepsilon_i(M, X)$ such that:
\begin{equation}
\sup_{x \in X} \|M(x)_i - C(M)(x)_i\|_\infty \leq \varepsilon_i(M, X)
\quad \forall i \in \{1, \dots, d_{\text{out}}\}
\label{eq:svnn_def}
\end{equation}
where $X \subset \mathbb{R}^{d_{\text{in}}}$ is the validated input domain
and $C(M)$ denotes the compiler's generated code executing on the target
platform's arithmetic (here, IEC~61131-3 REAL $\equiv$ IEEE~754 binary32).
\end{definition}

This definition captures the essential requirement: the compiler must
provide an \textit{a priori} guarantee---before the model ever runs on the
PLC---that the deployed code will not silently diverge from the trained
model. The error bound $\varepsilon_i$ may depend on the model's parameters,
the compiler's approximation choices (e.g., LUT resolution), and the input
domain, but it must be computable from these quantities alone.

\subsection{Sufficient Conditions for SVNN}
\label{sec:svnn-conditions}

We now establish the architectural conditions that enable SVNN compilation.
The conditions formalize why some architectures admit tight, composable
error bounds while others do not.

\vspace{4pt}
\noindent\textbf{Theorem 2 (SVNN Sufficient Conditions).}
\label{thm:svnn_conditions}
Let $\mathcal{A}$ be a feedforward neural network architecture. If $\mathcal{A}$
satisfies Conditions~1 and~2 below, then $\mathcal{A}$ is structurally
verifiable---for any instance $M \in \mathcal{A}$ of finite depth $L$, a
compiler adhering to the SVNN compilation discipline can compute, from the
model parameters alone, a finite per-output error bound. If $\mathcal{A}$
additionally satisfies the contractivity refinement (Condition~3), this bound
admits the closed, \textit{depth-independent} geometric-series form
$O(M_{\max} \cdot h^2 \cdot d_{\max} / (1-\gamma))$, where $M_{\max}$ is the
maximum second-derivative bound, $h$ is the LUT grid spacing, $d_{\max}$ the
maximum activation count per layer, and $\gamma < 1$ the contractivity
constant; absent Condition~3, the bound remains finite for every fixed $L$
but may grow with depth.

The conditions are designed to \textit{discriminate}: a standard MLP layer
$\mathbf{h} = \sigma(W\mathbf{x} + \mathbf{b})$ composes an affine transform
and a pointwise activation in one coupled operation, violating Condition~1.
This prevents per-operation error induction (the proof's structural induction
requires each node to be either purely linear or purely element-wise).
MLPs with SiLU or GELU additionally violate the Z3-verifiability of
Condition~2 (the transcendental $\exp$ is undecidable in Z3's NRA theory);
MLPs with ReLU lack a $C^2$ curvature parameter. KAN layers, in contrast,
decompose naturally at the IR level into separate MatMul, BsplineLUT,
StandardAct, and Add nodes---each satisfies one of the two pure types.
Conditions~1--2 thus characterize exactly the class that admits the
induction, and KAN belongs to it; standard MLPs do not.

\begin{enumerate}
    \item \textbf{Operation-Type Closure (Condition~1).} The architecture
    can be decomposed into a directed acyclic graph where every node is
    either (a) a linear transformation (matrix multiplication, vector
    addition) or (b) an element-wise univariate function. No node may
    mix these operation types. This decomposition is provided by the
    compiler's IR ({\S}\ref{sec:ir}); it guarantees that the per-node
    error induction in the proof below never encounters a coupled
    linear+nonlinear operation.

    \item \textbf{Univariate Curvature Bound (Condition~2).} Every element-wise
    univariate function $f: \mathbb{R} \to \mathbb{R}$ is $C^2$-continuous
    on the validated input domain $X$, and its second derivative satisfies
    $\sup_{x \in X} |f''(x)| \leq M_f < \infty$. The bound $M_f$ must be
    computable from the function's parameters alone. This enables the
    de~Boor LUT error bound $\varepsilon_f = M_f h^2/8$---a
    \textit{compiler-controllable} error: by increasing the LUT resolution
    (reducing $h$), the per-function error decreases quadratically.
    Condition~2 thus provides the compiler with a tunable guarantee knob
    that MLPs with ReLU (no $C^2$) or SiLU/GELU (transcendental,
    Z3-undecidable) lack.
\end{enumerate}

\noindent Conditions~1 and~2 are the \textit{sufficient conditions} for
structural verifiability. The following refinement governs how the bound
scales, not whether it exists:

\begin{enumerate}\setcounter{enumi}{2}
    \item \textbf{Contractive Composition (Condition~3, refinement).} The
    product of each element-wise layer's Lipschitz constant
    $L_f^{(\ell)} = \sup_x |f'(x)|$ with the subsequent linear layer's
    row-sum norm satisfies $L_f^{(\ell)} \cdot \|W^{(\ell+1)}\|_{1,\infty}
    \leq \gamma < 1$. When this holds, the bound converges to a
    depth-independent constant $O(1/(1-\gamma))$; when it does not, the
    bound remains finite for the specific deployment depth but grows as
    $O(\gamma^L)$. Condition~3 upgrades a per-depth certificate to a
    depth-independent one; it is \textit{not} required for SVNN membership.
\end{enumerate}

\vspace{4pt}
\noindent\textit{Proof.} The proof proceeds in four steps.

\vspace{2pt}
\noindent\textbf{Step 1 (Linear Layer Exactness).}
Let $y = Wx + b$ be a linear transformation with input $x$ that carries
error $\delta_x$ (i.e., the compiled input is $\tilde{x} = x + \delta_x$).
The compiled output is $\tilde{y} = W\tilde{x} + b = Wx + b + W\delta_x$.
Hence $\|\tilde{y} - y\|_\infty = \|W\delta_x\|_\infty \leq
\|W\|_{1,\infty} \cdot \|\delta_x\|_\infty$. A linear layer
introduces \textit{no new error sources}---it only amplifies or attenuates
existing error. Under sign-structural affine arithmetic ({\S}\ref{sec:da}), the
amplification factor exploits weight-matrix sign structure, yielding
$\|W\delta_x\|_\infty \leq \|W^+\delta_x^+ + W^-\delta_x^-\|_\infty$,
which is strictly tighter than the interval-arithmetic bound
$\|W\|_{1,\infty} \cdot \|\delta_x\|_\infty$ when $W$ contains both
positive and negative entries.

\vspace{2pt}
\noindent\textbf{Step 2 (Element-Wise Boundedness).}
Consider an element-wise univariate function $f$ applied to input $x$.
The compiled version uses a piecewise-linear LUT approximation
$\tilde{f}$ on a grid with spacing $h$. By the de~Boor--Cox recurrence
for B-spline approximation error~\cite{deBoor1978}, for any $C^2$
function:
\begin{equation}
\|f - \tilde{f}\|_\infty \leq \frac{M_f \cdot h^2}{8}
\label{eq:de_boor_general}
\end{equation}
where $M_f = \sup_x |f''(x)|$. Condition~2 guarantees that $M_f$ is
finite and computable. The LUT thus introduces fresh error at most
$\varepsilon_f = M_f h^2 / 8$ per function evaluation.

If the input carries error $\delta_x$, the mean value theorem gives:
\begin{equation}
|f(x + \delta_x) - \tilde{f}(x + \delta_x)| \leq
|f(x + \delta_x) - f(x)| + |f(x) - \tilde{f}(x + \delta_x)|
\leq L_f \cdot |\delta_x| + \varepsilon_f
\label{eq:element_error}
\end{equation}
where $L_f = \sup_x |f'(x)|$ is the Lipschitz constant of $f$.

\vspace{2pt}
\noindent\textbf{Step 3 (Finite-Depth Composition---Conditions 1 and 2 suffice).}
We now compose the analysis across $L$ layers by structural induction on
the layer index $\ell$. Let $\Delta^{(\ell)}$ denote the maximum per-output
error after layer $\ell$, with $\Delta^{(0)} = 0$ (input is exact).

For a KAN layer (element-wise B-spline activation followed by a linear
combination via weight matrix):
\begin{equation}
\Delta^{(\ell)} \leq L_f^{(\ell)} \cdot \|W^{(\ell)}\|_{1,\infty}
\cdot \Delta^{(\ell-1)} + \varepsilon_f^{(\ell)} \cdot d_{\ell-1}
\label{eq:layer_recurrence}
\end{equation}

The first term captures error amplification from the previous layer; the
second term captures fresh approximation error from $d_{\ell-1}$ parallel
element-wise function evaluations (one per input dimension).

Unrolling the recurrence across $L$ layers yields the total bound:
\begin{equation}
\Delta^{(L)} \leq \sum_{\ell=1}^{L} \varepsilon_f^{(\ell)} \cdot
d_{\ell-1} \cdot \prod_{j=\ell+1}^{L}
\bigl(L_f^{(j)} \cdot \|W^{(j)}\|_{1,\infty}\bigr)
\label{eq:total_bound}
\end{equation}

The critical role of Condition~1 is now apparent: it guarantees that every
node encountered by the induction is of a known type with a closed-form error
rule (Step~1 for linear nodes, Step~2 for element-wise nodes). The recurrence
Eq.~\ref{eq:layer_recurrence} applies \emph{only} because each node is
provably pure. In a standard MLP layer $\mathbf{h} = \sigma(W\mathbf{x} +
\mathbf{b})$, the affine transform and pointwise activation are coupled in a
single operation---there is no intermediate variable whose error can be
separately bounded, so the structural induction cannot be initiated. This is
why Sigmoid/Tanh MLPs, despite having bounded $M_2$ (they satisfy
Condition~2), are not SVNN: they fail Condition~1.

Under Conditions~1 and~2, every term in Eq.~\ref{eq:total_bound} is finite
and computable: $\varepsilon_f^{(\ell)} = M_f^{(\ell)} h^2/8$ from
Condition~2, $L_f^{(j)}$ finite on the compact domain $X$, and
$\|W^{(j)}\|_{1,\infty}$ finite by construction. Eq.~\ref{eq:total_bound}
is therefore a finite sum of finite terms, directly yielding a computable
design-time bound $\Delta^{(L)} < \infty$ for any fixed depth $L$---the
existence guarantee of Definition~\ref{def:svnn}. The bound is
\emph{compiler-controllable}: increasing the LUT resolution (reducing $h$)
monotonically decreases $\varepsilon_f$ and hence $\Delta^{(L)}$, giving the
compiler a tunable knob to meet any target safety requirement.

\vspace{2pt}
\noindent\textbf{Step 3b (Depth-Uniform Refinement---Condition 3).}
Condition~3 does not change \textit{whether} a bound exists; it controls
\textit{how the bound scales with depth}. When
$L_f^{(j)} \cdot \|W^{(j)}\|_{1,\infty} \leq \gamma < 1$, the product terms in
Eq.~\ref{eq:total_bound} decay geometrically, and the finite sum is majorized
by the convergent series
\begin{equation}
\Delta^{(L)} \leq \frac{\max_\ell(\varepsilon_f^{(\ell)} \cdot
d_{\ell-1})}{1 - \gamma}
= O\!\left(\frac{M_{\max} \cdot h^2 \cdot d_{\max}}{1-\gamma}\right)
\label{eq:convergent_bound}
\end{equation}
where $M_{\max} = \max_f M_f$ and $d_{\max} = \max_\ell d_{\ell-1}$
is the maximum number of element-wise activation functions per layer.
The bound follows by substituting $\varepsilon_f^{(\ell)} =
M_f^{(\ell)} h^2 / 8$ (Eq.~\ref{eq:de_boor_general}) into the
geometric series, yielding a per-layer factor of $M_{\max} h^2 d_{\max}$
multiplied by the geometric decay $1/(1-\gamma)$. This form is
\emph{depth-independent}: when $\gamma < 1$, the bound converges to a
constant as $L \to \infty$, admitting deployment on
arbitrarily deep networks. When Condition~3 does not hold ($\gamma \geq 1$),
Eq.~\ref{eq:total_bound} still evaluates to a finite number for the fixed
deployment depth, but the worst-case majorant grows as
$O(M_{\max} \cdot h^2 \cdot \gamma^L)$---still computable, only looser with
depth. Thus Condition~3 upgrades a per-depth guarantee into a depth-uniform
one; it is a refinement, not a membership requirement.

\vspace{2pt}
\noindent\textbf{Step 4 (Tightness and Practical Considerations).}
The $O(L \cdot M_{\max} \cdot h^2 \cdot d_{\max})$ bound is a \textit{worst-case}
guarantee. In practice, three factors make the bound substantially
tighter for the KAN architecture specifically:
\begin{enumerate}
    \item \textbf{Segment-Aware $M_2$ ({\S}\ref{sec:segment-aware}):}
          Rather than using the global $\sup |f''|$, we compute
          per-segment $M_2^{(k)}$ bounds, tightening the LUT error
          estimate by $6.0\times$ on average for KAN B-splines.
    \item \textbf{sign-structural affine arithmetic ({\S}\ref{sec:da}):}
          Linear layers propagate error via DA rather than IA,
          exploiting weight-matrix sign structure for an additional
          $3.1\times$ tightening.
    \item \textbf{Adaptive LUT Allocation ({\S}\ref{sec:adaptive-lut}):}
          Non-uniform knot placement allocates more LUT points to
          high-curvature regions, reducing the effective $h$ for
          the dominant error sources by up to $71.6\%$.
\end{enumerate}
Together, these refinements transform the asymptotic $O(L \cdot M_{\max}
\cdot h^2 \cdot d_{\max})$ bound into a practically useful guarantee. For the trained
$[28,16,4]$ KAN, the combined segment-aware DA safety factor is approximately
$11.9\times$ above the minimum required for classification correctness
({\S}\ref{sec:method}, Table~\ref{tab:da_bounds_summary}).

\vspace{4pt}
\begin{remark}[Near-Necessity of the SVNN Conditions]
\label{rem:near_necessity}
While Theorem~2 establishes sufficiency, the question of necessity is
equally important: are these conditions arbitrarily chosen, or do they
characterize a genuine structural boundary? Theorem~5
({\S}\ref{sec:svnn-necessity}) provides the formal answer:
violating Condition~1 makes tight bound computation NP-hard,
establishing that the SVNN conditions are not an arbitrary restriction
but the precise boundary at which polynomial-time compiler-computed
correctness becomes achievable. We briefly summarize the three
observations that motivate the formal proof.

\vspace{2pt}
\noindent\textit{(a) Condition~1 is structurally necessary for polynomial-time
bound computation.}
Any architecture that couples linear transforms with nonlinear activations
in a single operation---the standard MLP layer $\mathbf{h} = \sigma(W\mathbf{x}
+ \mathbf{b})$---eliminates the intermediate representation whose error the
structural induction bounds. Without this decomposition, error propagation
must model the coupled operation as a single black-box function, requiring
general-purpose LP/SMT solvers (CROWN, Marabou, Reluplex) whose cost grows
with network size. Condition~1 thus separates architectures admitting $O(L
\cdot d_{\max}^2)$ compiler-computed bounds from those requiring
exponential-time SMT case-splitting.

\vspace{2pt}
\noindent\textit{(b) Condition~2 separates $C^2$-bounded from transcendental.}
The de~Boor LUT error bound ($\varepsilon = M_2 h^2/8$) requires $C^2$
continuity. Activations without $C^2$ (ReLU: $f''$ undefined at 0) or with
unbounded curvature (SiLU: $\sup |\text{SiLU}''| = \infty$ on $\mathbb{R}$)
cannot use this bound. The $C^2$ requirement is nearly tight: it excludes
exactly the activations whose LUT error cannot be bounded by a finite,
parameter-computable constant. On the compact input domain $[-3,3]^{28}$,
SiLU's second derivative is finite ($\max \approx 0.21$)---but the key
distinction is \textit{computability}: $M_2^{\text{B-spline}}$ is computable
from the learned coefficients alone, whereas $\sup_{x\in[-3,3]}
|\text{SiLU}''(x)|$ requires numerical optimization over a transcendental
function---a different computational model.

\vspace{2pt}
\noindent\textit{(c) Richardson's Theorem~\cite{richardson1969decidability}
imposes a fundamental barrier.}
The equality of expressions involving $e^x$ is undecidable by
Richardson's theorem. Since SiLU contains $e^{-x}$, determining whether
$\text{SiLU}_{\text{LUT}}(x) = \text{SiLU}_{\text{FP32}}(x)$ for a given LUT
resolution is, in general, undecidable. KAN's B-spline activations are
piecewise-polynomial and avoid this barrier entirely. The SVNN conditions
are therefore not an arbitrary restriction---they characterize the precise
boundary at which compiler-computed correctness becomes \textit{decidable}.
\end{remark}

\vspace{4pt}
\noindent\textbf{Remark 2 (Relationship to Theorem~1).}
Theorem~1 ({\S}\ref{sec:method}) is the \textit{concrete instantiation}
of the SVNN framework for a specific architecture (2-layer KAN) and
compiler (NeuroPLC). Theorem~2 provides the \textit{general principle}:
any finite-depth architecture satisfying Conditions~1--2 admits a
correctness-preserving compiler, with Condition~3 sharpening the resulting
bound to a depth-uniform form. Theorem~1 proves that KAN satisfies the
conditions; Theorem~2 shows \textit{why} this is possible and characterizes
the broader class of architectures for which similar guarantees can be obtained.

\vspace{4pt}
\noindent\textbf{Remark 3 (Arithmetic Bounds vs.\ MILP-Based Verification).}
The SVNN arithmetic bound ($O(L \cdot M_{\max} \cdot h^2 \cdot d_{\max})$, Theorem~2)
and the PWA abstraction + MILP approach of Schwartz
et~al.~\cite{schwartz2026optimal} represent two complementary points on
the verification cost-precision spectrum. \textit{Arithmetic bounds}
(Level~2) are computable in closed form at compile time, require no
external solver, and provide a sound over-approximation for all inputs
in the validated domain---but they are inherently conservative (the
$\sim$11.9$\times$ safety factor for our trained KAN reflects this
conservatism). \textit{MILP-based verification} (Level~3) can provide
exact certification within the PWA abstraction error (typically
$<$0.1\% of the function range per unit), but requires solving an
MILP whose size grows with the number of PWA pieces---tractable for
KANs (each univariate function is abstracted independently) but
exponential for MLPs (multivariate activations require piecewise
decomposition of $\mathbb{R}^d$). The two approaches are synergistic:
arithmetic bounds serve as a fast, no-solver-required pre-certificate;
if the bound is within safety margins, no further verification is
needed. If tighter guarantees are required (e.g., for SIL~3+
certification), the SVNN decomposition enables targeted MILP
verification of safety-critical outputs. This two-tier architecture is
discussed further in {\S}\ref{sec:compositional}.



\vspace{4pt}
\noindent\textbf{Architecture Taxonomy Under the Compilable Frontier.}
Table~\ref{tab:svnn_taxonomy} maps major neural architecture families to the
SVNN conditions, providing a concrete instantiation of the compilable frontier
(Remark~\ref{rem:compilable_frontier}). The taxonomy reveals a stark divide:
architectures that decompose into pure linear $+$ element-wise operations
(Condition~1) and use $C^2$ parameter-computable activations (Condition~2)
occupy the \textit{interior}---their deployed behavior is certifiable at design
time. All architectures violating either condition lie on the
\textit{exterior}---they can be compiled and deployed, but correctness
guarantees are limited to test-set validation. The Contractivity column
(Condition~3) further distinguishes architectures whose error bounds are
depth-\textit{uniform} ($\gamma < 1$, $O(L)$ growth) from those whose bounds
are merely depth-\textit{finite} (computable but potentially growing with $L$).

\begin{table}[t]
\centering
\caption{Architecture Taxonomy Under the SVNN Compilable Frontier. Architectures on the interior admit compiler-computed design-time correctness certificates; architectures on the exterior require runtime testing.}
\label{tab:svnn_taxonomy}
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{@{}p{1.7cm}p{0.95cm}p{0.95cm}p{1.1cm}p{1.15cm}p{1.25cm}@{}}
\toprule
\textbf{Architecture} & \textbf{Cond.~1} & \textbf{Cond.~2} & \textbf{Contract.} & \textbf{Frontier} & \textbf{Z3-Verif.} \\
\midrule
KAN (B-spline) & \cmark & \cmark & \cmark$^{*}$ & Interior$^{\dag}$ & 512/512 \\
KAN (Chebyshev) & \cmark & \cmark & \cmark & Interior & Full$^{\ddag}$ \\
KAN (Wavelet) & \cmark & \cmark & $\sim$ & Interior & Full$^{\ddag}$ \\
\midrule
MLP + ReLU & \xmark & \xmark$^{\S}$ & N/A & Exterior & 0/16 \\
MLP + SiLU/GELU & \xmark & \xmark$^{\P}$ & N/A & Exterior & 0/16$^{\sharp}$ \\
MLP + Sigmoid & \xmark & \cmark & N/A & Exterior & -- \\
MLP + Tanh & \xmark & \cmark & N/A & Exterior & -- \\
\midrule
Transformer & \xmark & \xmark & N/A & Exterior & N/A \\
CNN (Conv2d) & \xmark & \xmark & N/A & Exterior & N/A \\
\bottomrule
\end{tabular}
\vspace{2pt}
\parbox{\linewidth}{\scriptsize
\cmark{} = condition satisfied; \xmark{} = condition violated; $\sim$ = depends on wavelet basis choice.\\
$^{*}$Contractivity verified for the trained KAN student $[28,16,4]$ under VRM-KD distillation; Tankman~\cite{tankman2026lipschitz} proves $P(\text{KAN}) \leq 1$ as a structural property for standard operation sets on $[0,1]$-bounded inputs.\\
$^{\dag}$The B-spline KAN is the only architecture in this taxonomy for which all three conditions are simultaneously satisfied \textit{and} empirically validated on physical hardware (TIA Portal V21, S7-1200).\\
$^{\ddag}$ChebyKAN: formally verified in Proposition~\ref{prop:chebykan-svnn} ({\S}\ref{sec:svnn-chebykan}); Z3 verifiability rate 496/512, 96.9\% (E54). WaveletKAN with piecewise-polynomial basis: full verifiability follows by the same structural induction as B-spline KAN (Theorem~2, Step~3); Z3 rate depends on wavelet family (Remark~4).\\
$^{\S}$ReLU is piecewise-linear ($C^0$); $f''$ is a Dirac delta at $x=0$, so the de~Boor $M_2 h^2/8$ bound does not apply. MILP-based verification (e.g., Reluplex) is possible but requires exponential case-splitting.\\
$^{\P}$SiLU contains $\sigma(x)=1/(1+e^{-x})$; $e^{-x}$ is transcendental and undecidable in Z3's NRA theory. On compact domain $[-3,3]$, $\sup|\text{SiLU}''| \approx 0.21$ is finite but requires numerical optimization---not computable from parameters alone.\\
$^{\sharp}$Confirmed empirically: MLP $[28,16,4]$ with SiLU, 0/16 components Z3-verifiable (E41).}
\end{table}

The taxonomy clarifies a structural fact with direct engineering consequences:
\textit{no} MLP variant---regardless of activation choice---crosses the
compilable frontier. ReLU MLPs violate Condition~2 (no $C^2$ curvature
parameter); SiLU/GELU MLPs violate Condition~2 (transcendental, Z3-undecidable)
\textit{and} Condition~1 (coupled linear+nonlinear); Sigmoid/Tanh MLPs satisfy
Condition~2 but violate Condition~1. The SVNN conditions together form a
\textbf{conjunction}---failure of either is sufficient to place an architecture
on the exterior. This is why the compiler's ability to provide guarantees is an
architectural question, not a compiler engineering question (Remark~%
\ref{rem:compilable_frontier}).

\subsection{KAN Satisfies the SVNN Conditions}
\label{sec:svnn-kan}

\begin{proposition}[KAN Is Structurally Verifiable]
\label{prop:kan-svnn}
Every Kolmogorov-Arnold Network (KAN) architecture, as defined by
Liu et~al.~\cite{liu2024kan}, satisfies the SVNN sufficient conditions
(Conditions~1 and~2) of Theorem~2 and is therefore an SVNN architecture.
The trained KAN student additionally satisfies the contractivity refinement
(Condition~3), placing it in the depth-uniform regime.
\end{proposition}

\vspace{4pt}
\noindent\textit{Proof.} We verify Conditions~1 and~2, which establish SVNN
membership, and then confirm the Condition~3 refinement:

\vspace{2pt}
\noindent\textit{Condition~1 (Operation-Type Closure).}
A KAN layer with $d_{\text{in}}$ inputs and $d_{\text{out}}$ outputs
decomposes naturally as:
\begin{equation}
\text{KANLayer}(x)_j = \underbrace{b(x_j)}_{\text{Base path}}
+ \sum_{i=1}^{d_{\text{in}}}
\underbrace{\phi_{j,i}(x_i)}_{\text{Spline path}}
\label{eq:kan_decomposition}
\end{equation}
where $b(\cdot)$ is a SiLU activation (element-wise univariate) and each
$\phi_{j,i}(\cdot)$ is a cubic B-spline (element-wise univariate).
The summation $\sum_{i=1}^{d_{\text{in}}} \phi_{j,i}(x_i)$ is a linear
combination, equivalent to a MatMul with an identity-projected weight
structure. The final addition $b(x_j) + \sum_i \phi_{j,i}(x_i)$ is
element-wise linear. Thus every KAN layer consists solely of
element-wise univariate functions and linear combinations---satisfying
Condition~1.

This decomposition is exactly the basis of NeuroPLC's IR design: each
KAN layer maps to a MatMul node, $d_{\text{in}} \times d_{\text{out}}$
BsplineLUT nodes, and $d_{\text{out}}$ Add nodes ({\S}\ref{sec:ir}).
The 6-op IR type system is not an arbitrary design choice---it is the
minimal closure required by the SVNN decomposition of KAN.

\vspace{2pt}
\noindent\textit{Condition~2 (Univariate Boundedness).}
Cubic B-splines ($k=3$) are $C^2$-continuous piecewise polynomial
functions. On the bounded input domain $X = [-3, 3]$ used by the KAN
student, the second derivative $\phi''(x)$ of each B-spline activation
is bounded. The bound $M_2 = \max_x |\phi''(x)|$ is computable directly
from the B-spline control points $\{w_c\}$ and knot vector $\{t_c\}$
via the derivative recurrence:
\begin{equation}
\phi''(x) = \sum_{c=0}^{G+k-2} w_c \cdot B''_{c,3}(x)
\label{eq:bspline_second_deriv}
\end{equation}
where $B''_{c,3}$ is computed from the B-spline basis derivative
recurrence~\cite{deBoor1978}. For the SiLU activation used in the base
path, $\text{SiLU}''(x) = \sigma(x)(2 - 2x\sigma(x) + x\sigma(x)
(1-\sigma(x)))$, which is bounded on any compact domain.

\vspace{4pt}
\noindent\textbf{Analytical $M_2$ and $L_B$ Upper Bounds.}
While $M_2 = \max_x |\phi''(x)|$ can be computed numerically from the B-spline
control points (as done in E11, yielding the model-specific value 0.177), we
further establish \textit{analytical} upper bounds that depend only on the
B-spline degree and knot spacing---providing a true a priori guarantee that
requires no forward pass or empirical measurement.

\textit{B-spline derivative bounds on uniform knots.}
For a cubic B-spline ($k=3$, order 4) defined on uniform knots with spacing $h$,
the basis function derivatives satisfy the following analytical bounds
(derived from the Cox-de~Boor recurrence~\cite{deBoor1978}):

\begin{equation}
\max_x |B'_{c,4}(x)| \leq \frac{3}{h},
\qquad
\max_x |B''_{c,4}(x)| \leq \frac{6}{h^2}
\label{eq:bspline_basis_bounds}
\end{equation}

For a B-spline curve $\phi(x) = \sum_c w_c B_{c,4}(x)$, at any $x$ at most 4
basis functions are non-zero (local support property). Consequently:

\begin{equation}
\max_x |\phi'(x)| \leq \frac{12 \cdot W_{\max}}{h},
\qquad
\max_x |\phi''(x)| \leq \frac{24 \cdot W_{\max}}{h^2}
\label{eq:bspline_curve_bounds}
\end{equation}

where $W_{\max} = \max_c |w_c|$ is the maximum absolute control-point value.
A tighter bound using control-point differences~\cite{deBoor1978} yields:

\begin{equation}
L_B^{\text{analytic}} \leq \frac{3}{h} \cdot \max_c |\Delta w_c|,
\qquad
M_2^{\text{analytic}} \leq \frac{6}{h^2} \cdot \max_c |\Delta^2 w_c|
\label{eq:bspline_diff_bounds}
\end{equation}

where $\Delta w_c = w_{c+1} - w_c$ and $\Delta^2 w_c = w_{c+1} - 2w_c + w_{c-1}$
are the first and second forward differences of the control-point sequence.

\textit{Application to KAN.}
For the KAN student $[28,16,4]$ with grid size $G=8$ on input domain $[-3,3]$,
the knot spacing is $h = 6/G = 0.75$. With control points typically bounded by
$W_{\max} \approx 0.3$ (enforceable via weight clipping during training),
Eq.~\ref{eq:bspline_curve_bounds} yields the a priori bounds
$L_B^{\text{analytic}} \leq 12 \times 0.3 / 0.75 = 4.8$ and
$M_2^{\text{analytic}} \leq 24 \times 0.3 / 0.5625 = 12.8$. The tighter
difference-based bounds (Eq.~\ref{eq:bspline_diff_bounds}) reduce these
to $L_B^{\text{analytic}} \leq 4 \times \max|\Delta w|$ and
$M_2^{\text{analytic}} \leq 10.67 \times \max|\Delta^2 w|$, both computable
from the control points alone before any forward pass.

\textit{Relationship to empirical values.}
The empirically measured values ($L_B = 0.65$, $M_2 = 0.177$ from E11) are
substantially tighter than the analytical worst-case bounds, because the KAN's
learned B-spline control points exhibit slow variation ($\max|\Delta w_c| \ll
W_{\max}$). The analytical bounds serve as the \textbf{architectural guarantee}
(proving that the architecture admits finite $M_2$ and $L_B$ a priori), while
the empirical measurements provide \textbf{model-specific refinement}
(tightening the bound post-training). Both are computed from the model
parameters alone---neither requires hardware execution. This two-tier approach
satisfies SVNN Condition~2: $M_f$ is analytically bounded by the architecture
and computable from parameters, with the analytical bound providing the
design-time certificate and the empirical refinement providing the deployment-time
tightening.

\vspace{2pt}
\noindent\textit{Condition~3 (Contractive Composition, refinement).}
Conditions~1 and~2 already establish that KAN is structurally verifiable
(Theorem~2, Step~3): the finite-depth error bound of
Eq.~\ref{eq:total_bound} exists and is computable from the control points
alone. We now show that the trained KAN \emph{additionally} satisfies the
contractivity refinement, which upgrades this guarantee to the depth-uniform
$O(L)$ form of Eq.~\ref{eq:convergent_bound}.
For the trained KAN student $[28,16,4]$, the control-point difference bound
(Eq.~\ref{eq:bspline_diff_bounds}) yields $L_B^{\text{analytic}} \leq 1.84$
using the trained control points (analytically computed, no forward pass).
The tighter empirical measurement gives $L_B = 0.65$ (maximum of $|\phi'(x)|$
across all 512 B-spline activations on a 1,000-point grid).
The weight matrix row-sum norms are
$\|W^{(0)}\|_{1,\infty} \leq 0.32$ and $\|W^{(1)}\|_{1,\infty} \leq 0.28$.
Even with the conservative analytical bound, the product satisfies
$L_B^{\text{analytic}} \cdot \|W^{(1)}\|_{1,\infty} \leq 1.84 \times 0.28
= 0.515 < 1$; with the empirical refinement, $L_B \cdot
\|W^{(1)}\|_{1,\infty} = 0.65 \times 0.28 = 0.182 \ll 1$.
Both satisfy the contractive condition, so the KAN student enjoys the
strongest, depth-uniform form of the SVNN guarantee. We emphasize, however,
that a KAN with larger weights violating $\gamma < 1$ would \emph{still} be
structurally verifiable by Conditions~1--2---it would simply fall back to the
finite depth-dependent bound. Contractivity is thus a property the trained
model happens to enjoy, not a precondition for its verifiability.

Beyond this empirical observation, Tankman~\cite{tankman2026lipschitz}
has recently provided a \textit{formal} characterization: for standard
KAN operation sets $\{+, -, \times, \sin, \cos\}$ on $[0,1]$-bounded
inputs, the layer-wise Lipschitz product satisfies
$P(\text{KAN}) \leq 1$, independent of input dimension $n$ and
network depth. This result establishes that the contractivity required
by Condition~3 is not merely an artifact of our specific trained model
or distillation procedure---it is a \textit{structural property} of
KAN architectures under standard operations. The SVNN framework thus
captures a formally verified architectural invariant rather than an
empirical coincidence. $\square$

\subsection{Standard MLPs Do Not Admit Tight SVNN Error Bounds}
\label{sec:svnn-mlp}

Having established that KAN architectures satisfy the SVNN conditions,
we now show that the most widely deployed alternative---the Multi-Layer
Perceptron (MLP) with standard activations---does not. This result
clarifies \textit{why} prior ML$\to$PLC tools provide no correctness
guarantees: the architecture itself precludes them.

\vspace{4pt}
\noindent\textbf{Proposition 1 (Standard MLPs Do Not Admit the SVNN Tight-Bound Structure).}
Let $\mathcal{M}$ be a feedforward MLP using ReLU, SiLU, Sigmoid, or Tanh
activations. Then $\mathcal{M}$ does not admit the compiler-computable
architecture-independent tight LUT error bound of Condition~2, for the
following reasons. This does \emph{not} mean MLPs are unverifiable in
general (tools such as Reluplex and $\alpha$-$\beta$-CROWN verify ReLU
networks)---it means the \emph{KAN-style} de~Boor curvature bound, which
is computable from parameters alone without knowing the input distribution,
does not apply to MLP activations.

\vspace{4pt}
\noindent\textit{Proof.} We analyze each activation function:

\vspace{2pt}
\noindent\textit{(a) SiLU / GELU (modern MLP standard): $f(x) = x \cdot \sigma(x)$.}
$\text{SiLU}''(x)$ is bounded and the function is $C^\infty$, so Condition~2
is formally satisfiable. However, SiLU contains $\sigma(x) = 1/(1+e^{-x})$,
which uses the transcendental function $e^{-x}$. Z3's Nonlinear Real
Arithmetic (NRA) theory is \emph{undecidable} for transcendental functions,
so per-function Z3 verification---the mechanism behind the compositional
certificate of {\S}\ref{sec:compositional}---yields no proof for any SiLU segment. Empirically:
MLP $[28,16,4]$ with SiLU activations achieves 0/16 Z3-verifiable
components (E41), confirming that SiLU-based MLPs are incompatible with the
compositional verification pipeline regardless of bound tightness.

\vspace{2pt}
\noindent\textit{(b) ReLU: $f(x) = \max(0, x)$.}
ReLU is piecewise-linear with kink at $x = 0$. A LUT whose knots include
$x = 0$ would reproduce ReLU exactly ($\varepsilon = 0$), so the
\emph{existence} of a valid bound is not the issue. The issue is that the
de~Boor curvature bound (Eq.~\ref{eq:de_boor_general}, $\varepsilon \leq
M_2 h^2/8$) requires $C^2$ continuity, which ReLU violates at $x=0$
($f''$ is a Dirac delta). Without the curvature characterization, the
compiler cannot compute an a priori LUT error from the architecture and
parameters alone: the error depends on where inputs fall relative to the
kink, which is input-distribution-dependent. Existing tools that verify
ReLU networks (Reluplex, $\alpha$-$\beta$-CROWN) instead enumerate
activation patterns via case-splitting, at exponential cost in the number
of ReLU units---not by a single closed-form parameter-computable bound.

\vspace{2pt}
\noindent\textit{(c) Sigmoid / Tanh.}
Both are $C^\infty$ with bounded $M_2$ ($\sup_x|\sigma''| \approx 0.096$,
$\sup_x|\tanh''| = 2/\sqrt{27} \approx 0.385$), so Condition~2 is
satisfiable. The bottleneck is the \emph{mixing} within each MLP layer:
each neuron combines linear and nonlinear operations in the same layer, while
SVNN Condition~1 requires layers to be \emph{either} purely linear
\emph{or} purely element-wise. Standard MLP layers violate Condition~1
(affine transform + pointwise activation in one step), breaking the clean
compositional error accounting that makes per-function Z3 verification
tractable.

\vspace{2pt}
\noindent\textbf{Consequence for MLP compilation.}
The fundamental issue is not that MLPs cannot be compiled---they can,
and \neuroplc does compile MLPs ({\S}\ref{sec:method}). The issue is
that the compiler cannot provide a \textit{tight, architecture-aware
error bound} for MLPs. Without a per-function $M_2$ bound (Condition~2
violation for ReLU) or with only coarse global Lipschitz bounds
(Sigmoid/Tanh), the best achievable MLP error bound is:
\begin{equation}
\Delta_{\text{MLP}}^{(L)} \leq O\bigl(L \cdot L_{\text{act}}^L \cdot
\|\mathbf{W}\|_{\text{prod}}\bigr)
\label{eq:mlp_bound}
\end{equation}
where $L_{\text{act}} = \max(1, L_{\text{ReLU}}, L_\sigma, L_{\tanh})$.
For $L_{\text{ReLU}} = 1$, the bound grows as $O(L \cdot
\|\mathbf{W}\|_{\text{prod}})$, which can easily exceed 1.0 for
networks deeper than 3--4 layers with typical weight magnitudes.
This makes the guarantee vacuous for all but the shallowest MLPs.

In contrast, the KAN error bound grows as $O(L \cdot M_{\max} \cdot
h^2 \cdot d_{\max})$, which is independent of weight magnitudes (the linear layers
are exact in DA) and decreases cubically with LUT resolution $h$.
This architectural distinction---not implementation quality---is why
\neuroplc provides guarantees for KAN that it cannot provide for MLP.
$\square$

% ============================================================================
% ChebyKAN Verification (Proposition 2)
% ============================================================================
\input{section_svnn_chebykan}

% ============================================================================
% Theorem 3 (Computational Necessity) + Theorem 4 (Generalization Bound)
% ============================================================================
\input{section_svnn_theorems}

\subsection{Implications for Compiler Design}
\label{sec:svnn-implications}

The SVNN framework has five concrete implications for how neural network
compilers targeting safety-critical embedded platforms should be designed:

\begin{enumerate}
    \item \textbf{Choose verifiable architectures.} Rather than attempting
          to verify the output of an arbitrary architecture---an approach
          that the floating-point verification community has shown to be
          fundamentally limited~\cite{szasz2025floating,munoz2025nofloat,
          mirman2025lipschitz}---compiler designers should target
          architectures that are \textit{structurally amenable} to
          verification. KAN is one such architecture; the SVNN conditions
          characterize the broader class.

    \item \textbf{Design the IR to reflect the architecture's
          decomposition.} NeuroPLC's 6-op IR is not a general-purpose
          neural network IR---it is purpose-built for the linear $+$
          element-wise decomposition that KAN architectures naturally
          exhibit. Each IR operation type corresponds to a proof obligation
          in the correctness theorem ({\S}\ref{sec:method}, Theorem~1).

    \item \textbf{Exploit architectural properties for tighter bounds.}
          The three refinements identified in Theorem~2 Step~4
          (Segment-Aware $M_2$, sign-structural affine arithmetic, Adaptive LUT
          Allocation) are possible \textit{because} KAN's architecture
          exposes the necessary structure: per-function second derivatives,
          per-layer weight-matrix sign patterns, and per-function curvature
          profiles. A general-purpose compiler targeting arbitrary
          architectures cannot exploit these properties because the
          architecture does not provide them.

    \item \textbf{Separate the correctness guarantee from the
          implementation.} Theorem~2 provides a \textit{design principle};
          Theorem~1 provides an \textit{instantiation} for a specific
          compiler and architecture. Future work can instantiate the
          SVNN framework for other architectures (e.g., transformer
          variants with element-wise attention, polynomial KANs) and
          other target platforms (e.g., FPGA via KANEL\'{E}~\cite{kanele2026},
          ARM Cortex-M via LUT-KAN~\cite{lutkan2024}), reusing the same
          proof structure with architecture-specific refinements.

    \item \textbf{Bridge design-time and mechanized verification.}
          The SVNN framework's arithmetic bounds (Level~2,
          {\S}\ref{sec:svnn-challenge}) are complementary to SMT/MILP-based
          formal verification (Level~3). The design-time bound provides a
          fast, automatically computable certificate at $O(L \cdot M_{\max}
          \cdot h^2 \cdot d_{\max})$ cost; when tighter guarantees are required (e.g., for
          SIL~3+ certification under IEC~61508), the SVNN decomposition
          enables tractable MILP encoding because each univariate function
          can be abstracted independently as a piecewise-affine (PWA)
          function, as demonstrated by Schwartz et~al.~\cite{schwartz2026optimal}.
          This two-tier approach---fast arithmetic screening at compile time,
          followed by targeted MILP verification of safety-critical paths---
          is uniquely enabled by the SVNN architecture. Standard MLPs do not
          admit this decomposition: their multivariate ReLU activations
          require exponentially many PWA pieces to abstract, making MILP-based
          verification intractable beyond trivial depths~\cite{schwartz2026optimal}.
\end{enumerate}

These implications position \neuroplc not merely as a tool for compiling
KANs to PLCs, but as the first instance of a broader compiler design
paradigm: \textit{structurally verifiable compilation}, where the compiler's
ability to provide correctness guarantees is achieved by targeting
architectures that are designed to be verifiable.
